Scheduling and Vectorization for MPC

### 2022-03-09 paper_SIMD.tex
removed section 3.2 from main latex file

\subsection{Problem Statement}
\label{sec:problem}

\ana{Ishaq? Basically, define sequential schedule, then define an equivalent parallel schedule. A parallel schedule is equivalent if it preserves def-use relations in sequential schedule, or in other words, schedules def ahead of the use. Problem is to minimize cost of Parallel schedule.}

\ana{Commenting out this section to have paper compile on Github. Will revert!}
%\begin{comment}

At the lowest level, we have two types of MPC instructions 1) local/non-interactive instruction (i.e. \add) and 2) remote/interactive instruction (i.e. \mul). Let the cost of a single \add instruction be given by a monotonically decreasing $f(n)$, where the argument $n$ is the number of \add instructions being executed in parallel. Similarly the cost of a single \mul is given by a monotonically decreasing function $g(n)$.

Given a serial schedule (a linear graph) of an MPC program i.e. a sequence of instructions $\mathcal{S}:=(S_1, \dots, S_n)$, where $S_i \in \{\text{\add, \mul}\}, 1 \leq i \leq n$, and a def-use dependency graph $G(V, E)$ corresponding to $\mathcal{S}$, our task is to construct a parallel schedule (another linear graph) $\mathcal{P} := (P_1, \dots, P_n)$ observing the following conditions:

\begin{enumerate}
    \item Multiple, not necessarily continuous instructions of the same type (i.e. either \add or \mul) from $\mathcal{S}$ can be grouped into a single $P_i$  However, all such instructions must be of the same type (either \add or \mul).
    \item Def-use dependencies of the graph $G(V, E)$ are should be preserved i.e. if instructions $S_i, S_j, i < j$ are a def-use (an edge exists from $S_i$ to $S_j$ in $G$), then they can only be mapped to $P_{i'}, P_{j'}, i' < j'$.
\end{enumerate}

Our goal is to construct minimize the height of the graph $\mathcal{P}$. Indeed, a graph $\mathcal{P}$ with minimum height will maximize parallelization.

\paragraph{Correctness} Correctness of $\mathcal{P}$ is guaranteed by definition. Since def-use dependencies are preserved, the function (being computed) remains the same.

\paragraph{Cost Comparison} For the sequential schedule $\mathcal{S}$ consisting of $L$ local and $R$ remote instructions, the total cost is $\mathit{cost}(\mathcal{S}) = L \cdot f(1) + R \cdot g(1)$. In the extreme case where all $L$ and all $R$ instructions can be parallelized, the cost of $\mathcal{P}$ is $\mathit{cost}(\mathcal{P}) = L \cdot f(L) + R \cdot g(R)$. Since both $f$ and $g$ are monotonically decreasing, $\mathit{cost}(\mathcal{P}) < \mathit{cost}(\mathcal{S})$. Cost of all other parallel schedules lies between the extremes of $\mathit{cost}(\mathcal{S})$ and $\mathit{cost}(\mathcal{P})$.

Note that we use an MPC-Source control flow graph (CFG) $G'(V', E')$ along with def-use graph $G(V,E)$ to construct $\mathcal{P}$. We consider a linearized MPC schedule $\mathcal{S}$ above for ease of exposition. The argument becomes slightly more involved when dealing with a graph $G'$ that may contain cycles.

% Given a def-use dependency graph $G(V, E)$ that is constructed from for the input program. Each vertex/node in this graph represents an operation/instruction in the program and an edge represents dependence i.e. an edge from $V_1$ to $V_2$ means the result of operation in $V_1$ is used in the operation in $V_2$. To build a parallel schedule, we translate this graph $G(V, E)$ into another dependency graph $G'(V', E')$ such that:

% \begin{enumerate}
%     \item We collapse multiple vertices of the same type (i.e. same operation) in $V$ into a single vertex in $V'$.
%     \item dependencies are not broken i.e. If there is an edge from $V_1$ to $V_2$ ($V_1, V_2 \in V$), and we put them in vertices $V'_1, V'_2 \in G'$ respectively, then an edge must exist from $V'_1$ to $V'_2$.
% \end{enumerate}

% We want to build a graph $G'$ with fewest possible vertices since fewer vertices means more parallelization.

%\ishaq{TODO: @Ana: I am still confused about what prevents a reviewer from saying, "you argue correctness of the schedule from unrolled MPC, you don't argue it for the pre-unrolled CFG that you use". Please comment.} \ana{The correctness proof answer this question. The unrolled schedule is the \emph{concretization} of MPC source. We define (1) the obvious $\gamma$, (2) we define partial order over MPC source codes (this is our abstract domain $A$), and (3) a partial order over unrolled schedules. Then we prove two theorems that $A \le A' \Rightarrow \gamma(A) \subseteq \gamma(A')$ and that the MPC source $\le$ Vectorized MPC source.}

%\end{comment}

\subsection{Scheduling is NP-hard}
\label{sec:np}

\ana{Commented out this section to have paper compile on Github. Will revert!}

\begin{comment}
\ishaq{The follow text is all from Linsay's thesis}.
\brevised{
We consider two operations, call them $A$ and $M$. $A$ and $M$ are two abstract MPC instruction, 
but as an example, $A$ stands for the ADD MPC instruction 
and $M$ stands for the MUL instruction. Each instruction in the program is either an $A$-instruction or
an $M$-instruction. In order to benefit from parallelization/amortization, we must schedule two or more 
$A$-instructions in the same parallel node (or two or more $M$-instructions in the same parallel node). 
Scheduling $A$-instructions in parallel with $M$-instruction does not benefit from amortization.
It incurs the exact same cost as scheduling the $A$-instructions in a node $P_A$, scheduling the $M$-instructions 
in a node $P_M$, and having $P_A$ precede $P_M$ in the parallel schedule. This is the difference 
between classical scheduling, as studied in parallel computing, and MPC scheduling.

In our theoretical treatment, we make the following assumptions:

\item $A$ and $M$ are of equal cost, 1 unit.
\item There is unlimited bandwidth---i.e., a single $A$-instruction (or $M$-instruction) costs as much as $N$ amortized $A$-instructions 
(or $M$-instructions), namely 1 unit.

%  Furthermore the run time cost of 
% computing operations of the same type remains the same regardless of the number of operations that are run.
% Roughly speaking, 1 A operation costs the same as 100 A operations and 1000 A operations.

Consider a loop body that consists of $n$ sequences: $S_1$, ... $S_n$ of $A$ and $M$ instructions. 
More precisely, the loop body is such that its instructions can be grouped into such sequences. 
$S_1$, ... $S_n$ can execute in parallel, however, all instructions within a sequence must 
execute sequentially. For example, consider the three sequences (the right arrow indicates a \emph{dependence},
meaning that the source node must execute before the target node): 
\begin{enumerate}
\item $A \rightarrow M \rightarrow A$
\item $A \rightarrow A \rightarrow A$
\item $M \rightarrow A \rightarrow M$
\end{enumerate}

A \emph{schedule} $P: P_1 \rightarrow P_2 \dots \rightarrow P_k$ is such that for each sequence 
$S_i$ in the set, if $S_i[k]$ precedes $S_i[k']$ in $S_i$ then $S_i[k]$ is scheduled in node $P_l$, $S_i[k]$ 
is scheduled in node $P_{l'}$, and $P_l$ precedes $P_{l'}$ in $P$.

The cost of a schedule $P$ is

\begin{equation}
\mathit{cost}(P) = \sum_{i=1}^k \mathit{cost}(P_i)
\end{equation}

where $\mathit{cost}(P_i) = 1$ if $P_i$ consists of $A$-instructions only, or $M$-instructions only, 
and $\mathit{cost}(P_i) = 2$ if $P_i$ mixes $A$-instructions and $M$-instructions.

The problem is to find a schedule $P$ with \emph{minimal cost}. For example, 
a schedule with minimal cost for the sequences above is
\[ A(1), A(2) \rightarrow M(1), A(2), M(3) \rightarrow A(1), A(2), A(3) \rightarrow M(3) \]
(The parentheses above indicate the sequence where the instruction comes from: (1), (2), or (3).)
The cost of this schedule is 5.

The problem of finding a schedule $P$ with a minimal $cost(P)$ for a given loop body has been shown
to be an NP-Hard problem, as it can be reduced to the problem of finding a \emph{shortest common supersequence}, 
a known NP-Hard problem\cite{Maier1978},\cite{Vazirani2010}. The shortest common supersequence problem is as follows: 
given two or more sequences find the the shortest sequence that contains all of the original sequences. This can be solved
in $O(n^k)$ time, where $n$ is the cardinality of the longest sequence and $k$ is the 
number of sequences. For our problem $n$ is the maximum length of a node and $k$ is the number
of total number of nodes. %\ana{I think it is clear without additional explanation...}
%\ana{Also, add the full statement of the problem here. Now it's not clear what's NP-hard, what's n, what's k, etc.?}
%\lindsey{not exactly sure what need to go it there, but does that make it clearer?}

To see the reduction, 
suppose $P$ is a schedule with minimal cost (computed by a black-box algorithm). 
We can derive a schedule $P'$ with the same cost as $P$, by mapping each mixed node $P_i \in P$ 
to two consecutive nodes in $P'$: an $A$-instruction node followed by an $M$-instruction node.
Clearly, $P'$, which now is a sequence of $A$'s and $M$'s, is a supersequence of each sequence 
$S_i$, i.e., $P'$ is a common supersequence 
of $S_1 \dots S_n$. It is also a shortest common supersequence. (To see this, suppose, a 
shorter common supersequence, $P''$, exists. $P''$ is a schedule of $S_1 \dots S_n$
and the $\mathit{cost}(P'')$ equals the length of $P''$. Since $P'$ is longer than
$P''$, and $\mathit{cost}(P') = \mathit{cost}(P)$, that means $\mathit{cost}(P'') < \mathit{cost}(P)$, 
which is a contradiction since $P$ is a schedule with minimal cost.)
}
\end{comment}

>>>>

\input{prob_stmt_np_hardness.tex}
